---
title: "Ubisoft Cybersecurity Analysis"
output: html_notebook
---

```{r}
num_samples <- 50
```

Load required packages.

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(tidyr))
```

Load NLP requirements.

```{r}
reticulate::use_condaenv("textrpp_condaenv")

library(text)

textrpp_initialize(condaenv = "textrpp_condaenv", prompt = FALSE)
```

This is an example using [Ubisoft](https://github.com/humbertcostas/vulnmine) data. 

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 
```{r}
# encoded UCS-2 LE with BOM
ubisoft_url <- "https://github.com/humbertcostas/vulnmine/raw/master/data/csv/v_GS_ADD_REMOVE_PROGRAMS.csv"
df_ubi_raw <- read.csv2(file = ubisoft_url, as.is = T, sep = "|", 
                        fileEncoding = "UCS-2", encoding = "UTF-16")
df_ubi_raw$id <- 1:nrow(df_ubi_raw)
```

Group inventory and get unique values for vendor, product and version:

```{r}
df_ubi <- df_ubi_raw %>% 
          select(Publisher0, DisplayName0, Version0) %>%
          rename("vendor" = "Publisher0", 
                 "product" = "DisplayName0", 
                 "version" = "Version0") %>% 
          group_by(vendor, product, version) %>% 
          summarise(n = n(), .groups = 'drop') %>%
          ungroup() %>%
          arrange(desc(n)) %>%
          mutate(id = seq_along(n)) %>%
          select(id, vendor, product, version, n)
df_ubi$id <- 1:nrow(df_ubi)

df_ubi_sample <- head(df_ubi, num_samples)

kbl(df_ubi_sample) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

Create a clean **title** string as new column using strings from vendor, product and version.

```{r}
df <- mitre::cpe_sccm_inventory(df_sccm = df_ubi, verbose = T)
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.
```{r}
predict_cpe <- function(df_inventory = mitre::getInventory(), 
                        model_name = "Neurona/cpener-test", 
                        hastitle = FALSE) {
  # Function for processing NER output
  embed2cpener <- function(df_ner = data.frame()) {
    df_cpe <- as.data.frame(df_ner$x_NER)
    df_cpe$entity <- stringr::str_replace_all(string = df_cpe$entity,
                                              pattern = "^[BIOLU]\\-(cpe.+)$",
                                              replacement = "\\1")
    df_cpe <- inner_join(x = df_cpe %>% 
                             group_by(entity) %>% 
                             summarise(score = mean(score)),
                         y = df_cpe %>% 
                             select(entity, word) %>% 
                             group_by(entity) %>% 
                             mutate(word = paste(word, collapse = " ")) %>% 
                             unique() %>% 
                             as.data.frame(),
                         by = "entity")
    df_cpe$word <- stringr::str_replace_all(string = df_cpe$word, pattern = "\\s##", replacement = "")
    df_cpe$word <- stringr::str_replace_all(string = df_cpe$word, pattern = "^\\s*##", replacement = "")

    return(df_cpe)
  }

  # Function for CPE creation
  cpener2cpe23 <- function(df_ner = data.frame()) {
    part <- "a"
    vendor <- ifelse(test = "cpe_vendor" %in% df_ner$entity,
                     yes = df_ner$word[df_ner$entity == "cpe_vendor"],
                     no = "*")
    product <- ifelse(test = "cpe_product" %in% df_ner$entity,
                      yes = df_ner$word[df_ner$entity == "cpe_product"],
                      no = "*")
    version <- ifelse(test = "cpe_version" %in% df_ner$entity,
                      yes = df_ner$word[df_ner$entity == "cpe_version"],
                      no = "*")
    vendor <- stringr::str_replace_all(vendor, " ", "_")
    product <- stringr::str_replace_all(product, " ", "_")
    version <- stringr::str_replace_all(version, "(\\d) (\\d)", "\\1\\.\\2")
    version <- stringr::str_replace_all(version, "(\\d) (\\d)", "\\1\\.\\2")
    if (vendor == "*") vendor <- product
    if (product == "*") product <- vendor
    cpe <- paste("cpe", "2.3", part, vendor, product, version, "*:*:*:*:*:*:*", sep = ":")
    cpe <- stringr::str_replace_all(cpe, "_([[:punct:]])_", "\\1")
    return(cpe)
  }

  if (!hastitle) {
    sw_title <- paste(mitre::cpe_wfn_vendor(df_inventory$vendor),
                      mitre::cpe_wfn_product(df_inventory$name), sep = " ")
    df_pred <- data.frame(title = sapply(sw_title,
                                         function(x)
                                           paste(unique(unlist(strsplit(x, " "))), collapse = " ")),
                          cpe = rep(NA, length(sw_title)))
    df_pred$title <- paste(df_pred$title, df_inventory$version)
  } else {
    df_pred <- df_inventory
  }

  df_inventory$cpe <- sapply(df_pred$title,
                             function(x) {
                               cpener2cpe23(embed2cpener(textNER(x = x, 
                                                                 model = model_name, 
                                                                 device = "gpu")))
                             }) %>% as.character()
  
  return(df_inventory)
}


```


```{r}
# df_inventory <- df %>% sample_n(5)

df_inventory <- df[df$id %in% df_ubi_sample$id,]

```

```{r}
df_inventory <- predict_cpe(df_inventory = df_inventory, hastitle = TRUE)

kbl(df_inventory) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r}
# replace cpe version (ner prediction) by orginal version
df_inventory <- df_inventory %>% 
                separate(col = cpe , sep = ":", extra = "merge", 
                         into = c("std", "v", "part", "vend", "prod", "vers", "tail")) %>%
                mutate(cpe = paste(std, v, part, vend, prod, version, tail, sep = ":")) %>%
                select(names(df_inventory))
```

## Evaluate predictions

### Prepare Official CPE tables

Prepare Official CPE vendor/product tables:  
 - cpes_vp: group by vendor AND product  
 - cpes_vend: group by vendor  
 - cpes_prod: group by prod  

```{r}
cpes <- mitre::cpe_latest_data(remote = T)

cpes_vp <- mitre::getCPEstats(cpes, only_vendor = FALSE)
cpes_vend <- mitre::getCPEstats(cpes)
cpes_prod <- cpes %>% group_by(product) %>% summarise(n = n(), .groups = "drop") %>% arrange(desc(n))
```

### Prepare features for evaluation

Prepare data.frame for evaluation and candidates selection with columns:  
 - vendor, product, version: Original values  
 - ner_vendor, ner_product, ner_version: Predicted values with NER model  
 - ner_cpelite: concatenate ner_vendor:ner_product  

```{r}
# prepare data.frame for evaluation and candidates selection
df_eval <- df_inventory %>% 
           separate(col = cpe , sep = ":", extra = "merge", 
                    into = c("std", "v", "part", "ner_vend", "ner_prod", "ner_vers", "tail")) %>%
           select(id, vendor, product, version, ner_vend, ner_prod, ner_vers) %>%
           mutate(ner_cpelite = paste(ner_vend, ner_prod, sep = ":"))
```

### Evaluations tests

Begin detecting vendor perfect match or close similarity with official CPE data.
For those vendors try to match the product with each vendors official products.

```{r}
df_match <- data.frame(id = integer(),
                       vd_vendor = character(),
                       vd_product = character(),
                       vd_match_type = character(),
                       vd_score = numeric(), stringsAsFactors = FALSE)
```

#### TEST M1: Perfect match for predicted vendor AND product

```{r}
# Perfect match with vendor using inner_join
df_test <- inner_join(df_eval %>% select(id, ner_vend, ner_prod),
                       cpes_vp %>% select(vendor, product), 
                       by = c("ner_vend" = "vendor", "ner_prod" = "product")) %>%
            rename("vd_vendor" = "ner_vend", "vd_product" = "ner_prod") %>%
            select(id, vd_vendor, vd_product)
df_test$vd_match_type <- rep("M1", nrow(df_test))
df_test$vd_score <- rep(1.0, nrow(df_test))
df_match <- bind_rows(df_match, df_test)

df_eval <- df_eval[!(df_eval$id %in% df_match$id), ]
```

#### TEST M1A: t(M1) AND predicted product IN CPEs (maybe from other vendor)

```{r}
df_test <- inner_join(df_eval %>% select(id, ner_vend, ner_prod),
                       cpes_vend %>% select(vendor), 
                       by = c("ner_vend" = "vendor")) %>%
           inner_join(cpes_prod %>% select(product), by = c("ner_prod" = "product")) %>%
           rename("vd_vendor" = "ner_vend", "vd_product" = "ner_prod") %>%
           select(id, vd_vendor, vd_product)
df_test$vd_match_type <- rep("M1A", nrow(df_test))
df_test$vd_score <- rep(.95, nrow(df_test))

df_match <- bind_rows(df_match, df_test)
df_eval <- df_eval[!(df_eval$id %in% df_match$id), ]
```

#### TEST M1B: t(M1) AND similar product IN CPEs (from same vendor)

```{r}
df_test <- inner_join(df_eval %>% select(id, ner_vend, ner_prod),
                       cpes_vp %>% select(vendor, product), 
                       by = c("ner_vend" = "vendor")) 
df_test$ner_prod <- stringr::str_replace_all(df_test$ner_prod, "_", " ")
df_test$product <- stringr::str_replace_all(df_test$product, "_", " ")
df_test$vd_match_type <- rep("M1B", nrow(df_test))
df_test$vd_score <- stringdist::stringsim(df_test$ner_prod, df_test$product, method = "jw")
# df_test$vd_score <- stringdist::stringsim(df_test$ner_prod, df_test$product, method = "osa")
# df_test$vd_score <- stringdist::stringsim(df_test$ner_prod, df_test$product, method = "qgram")

df_test <- inner_join(df_test %>% group_by(id) %>% summarise(vd_score = max(vd_score), .groups = "drop"),
                        df_test, by = c("id" = "id", "vd_score" = "vd_score")) %>%
           select(names(df_test)) %>% 
           filter(vd_score >= 0.8)
df_test$ner_prod <- stringr::str_replace_all(df_test$ner_prod, " ", "_")
df_test$product <- stringr::str_replace_all(df_test$product, " ", "_")

df_test <- df_test %>% 
           rename("vd_vendor" = "ner_vend", "vd_product" = "product") %>%
           select(names(df_match))

df_match <- bind_rows(df_match, df_test)
df_eval <- df_eval[!(df_eval$id %in% df_match$id), ]
```

#### TEST M1C: t(M1) AND NOT t(M1A, M1B) similar product IN CPEs (maybe from other vendor)



----------------

```{r}
vend_eval <- stringdist::stringsimmatrix(df_eval$ner_vend, cpes_vend$vendor)
vend_eval <- as.data.frame.matrix(vend_eval)
names(vend_eval) <- cpes_vend$vendor

df_eval$cpe_vend <- names(vend_eval)[apply(vend_eval, 1, which.max)]
df_eval$cpe_vend_dist <- apply(vend_eval, 1, max)
```

```{r}
# prod_eval <- stringdist::stringsimmatrix(df_eval$ner_prod, cpes_prod$product)
# prod_eval <- as.data.frame.matrix(prod_eval)
# names(prod_eval) <- cpes_prod$product
# 
# df_eval$cpe_prod <- names(prod_eval)[apply(prod_eval, 1, which.max)]
# df_eval$cpe_prod_dist <- apply(prod_eval, 1, max)
```

```{r}
cand_vend <- df_eval[((df_eval$cpe_prod_dist == 1) & (df_eval$cpe_vend_dist < 1)), c("id", "vend", "cpe_prod")]
cand_vend <- left_join(cand_vend %>% rename("product" = "cpe_prod"), 
                       cpes_vp %>% select(vendor, product), by = "product") %>%
             select(id, vendor) %>% rename("cand_vend" = "vendor")
df_eval <- left_join(df_eval, cand_vend, by = "id")
df_eval$cand_vend[is.na(df_eval$cand_vend)] <- ""
df_eval$cand_vend_dist <- apply(df_eval, 1, function(x) stringdist::stringsim(x["vend"], x["cand_vend"]))
```

```{r}
cand_prod <- df_eval[((df_eval$cpe_vend_dist == 1) & (df_eval$cpe_prod_dist < 1)), c("id", "prod", "cpe_vend")]
cand_prod <- left_join(cand_prod %>% rename("vendor" = "cpe_vend"), 
                       cpes_vp %>% select(vendor, product), by = "vendor") %>%
             # select(id, product) %>% 
             rename("cand_prod" = "product")
cand_prod$cand_prod_dist <- apply(cand_prod, 1, function(x) stringdist::stringsim(x["prod"], x["cand_prod"]))

cand_prod <- left_join(cand_prod %>% group_by(id) %>% summarise(cand_prod_dist = max(cand_prod_dist)),
                cand_prod %>% select(-prod, -vendor), by = (c("id", "cand_prod_dist"))) %>% 
             group_by(id, cand_prod_dist) %>% 
             summarise(cand_prod = paste(cand_prod, collapse = ":::"), .groups = "drop") %>%
             select(id, cand_prod, cand_prod_dist)


df_eval <- left_join(df_eval, cand_prod, by = "id")
df_eval$cand_prod[is.na(df_eval$cand_prod)] <- ""
df_eval$cand_prod_dist[is.na(df_eval$cand_prod_dist)] <- 0

```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
