---
title: "R Notebook"
output: html_notebook
---

```{r}
suppressPackageStartupMessages(library(dplyr))
```

This is an example using [Ubisoft](https://github.com/humbertcostas/vulnmine) data. 

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

```{r}
ubisoft_url <- "https://github.com/humbertcostas/vulnmine/raw/master/data/csv/v_GS_ADD_REMOVE_PROGRAMS.csv"
# encoded UCS-2 LE with BOM -> converted with notepad and removed noisy row 20045
# t <- "../extdata/ubisoft.csv"
# download.file(ubisoft_url, t)

df_ubi_raw <- read.csv("../extdata/ubisoft3.csv", header = T, sep = "|", fill = T)

df_ubi <- df_ubi_raw %>% select(Publisher0, DisplayName0, Version0)
names(df_ubi) <- c("vendor", "name", "version")

df <- mitre::cpe_local_inventory(df = df_ubi)

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
