---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mitre

<!-- badges: start -->
[![Travis build status](https://www.travis-ci.com/motherhack3r/mitre.svg?branch=master)](https://www.travis-ci.com/motherhack3r/mitre)
[![CRAN status](https://www.r-pkg.org/badges/version/mitre)](https://cran.r-project.org/package=mitre)
[![](https://cranlogs.r-pkg.org/badges/grand-total/mitre)](https://cran.r-project.org/package=mitre)
[![](https://www.rdocumentation.org/badges/version/mitre)](https://www.rdocumentation.org/packages/mitre)
<!-- badges: end -->

mitre package is designed to provide easy access to cybersecurity data standards.
You can expect sample data frames for every standard object.
It provide a directed graph with all relationships for deep exploratory analysis.
Using scripts in data-raw folder you can build the data sets using the latest 
public source files.

## Installation

You can install the released version of mitre from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("mitre")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("motherhack3r/mitre")
```
## Explore cybersecurity standards

This is a basic example which shows you how to view shield tactics ids and names:

```{r example}
library(mitre)
shield <- mitre::shield.tactics
shield[, c("id", "name")]
```

This example shows the distribution of vulnerability risk:

```{r example2}
hist(mitre::cve.nist$cvss3.score, 
     main = "CVE risk distribution", xlab = "cvss3")
```


## Standards network

This code shows you the type of nodes for each standard:

```{r example3}
mitrenet <- mitre::build_network(as_igraph = FALSE)
table(mitrenet$nodes$type, mitrenet$nodes$group)
```

And the type of relationships:
```{r example4}
table(mitrenet$edges$label)
```


## Code of conduct

Please note that this project is released with a [Contributor Code of Conduct](https://pkgdown.r-lib.org/CODE_OF_CONDUCT.html). By participating in this project you agree to abide by its terms.
