---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mitre

<!-- badges: start -->

[![Travis build status](https://travis-ci.com/humbertcostas/mitre.svg?branch=master)](https://travis-ci.com/humbertcostas/mitre)
<!-- badges: end -->

mitre package is designed to provide easy access to cybersecurity data starndards.
You can expect functions to get data frames for every standard object.
For deep exploratory, it provide visNetwork objects for graph and relationship
analysis.

## Installation

You can install the released version of mitre from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("mitre")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("humbertcostas/mitre")
```
## Basic usage

This is a basic example which shows you how to view shield tactics ids and names:

```{r example}
library(mitre)
tactics <- getShieldTactics()
tactics[, 1:2]
```

## Network visualization

This is a example which shows you how to visualize shield network:

```{r example2}
# library(mitre)
# shieldnet <- getShieldNetwork()
# shieldnet
```

For more advanced visualizations with shiny, check vignettes.

## Code of conduct

Please note that this project is released with a [Contributor Code of Conduct](https://pkgdown.r-lib.org/CODE_OF_CONDUCT.html). By participating in this project you agree to abide by its terms.
