---
title: "Mitre"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mitre}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

You can explore MITRE Shield data frames as you usually analyze data, but
the power reside in relations. Let's try to visualize the hole graph with 
a ShinyApp... let's call it: MITRExplorer.
  
# MITRExplorer
This is as Proof Of Concept project to provide an exploration tool for 
MITRE data sets.

Try to build your own MITRExplorer app using the following code for ui and server.

## Shinyapp demo
```{r shiny_demo}
# library(shiny)
# library(visNetwork)
# library(mitre)
# 
# mitredata <- mitre::getLatestDataSet()
# shieldnet <- mitredata$standards$shield$shieldnet
# 
# # Define UI for application that draws a histogram
# ui <- shinyUI(fluidPage(
#    # Application title
#    titlePanel("MITRE Shield Explorer"),
#    # Show a plot of the generated distribution
#    mainPanel(
#        visNetworkOutput("shieldnetwork")
#    )
# ))
# 
# ## server.R
# server <- shinyServer(function(input, output) {
#    output$shieldnetwork <- renderVisNetwork({
#       ggnet <- visNetwork(nodes = shieldnet$nodes,
#                           edges = shieldnet$edges)
#       ggnet %>%
#          visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) %>%
#          visLayout(randomSeed = 123)
#    })
# })
# 
# shiny::shinyApp(ui, server)
```

Check this app running in [here](https://security.shinyapps.io/mitreshield/) as free shinyapps.io service.

